---
title: "GSVA BioDomain for mutil Group"
author: "Ximing Ran"
date: "2025-04-09"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: false
  html_document:
    # code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js

---

```{r setup, include=FALSE}
# load libraries
library(tidyverse)
library(knitr)
set.seed(2024)

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.path = "./results/01-BioDomain/Analysis_figure/"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
knitr::kable(head(mtcars[, 1:4]), "simple")
```

```{r}
library(tibble)
library(tidyr)
library(dplyr)
library(rtracklayer)
```



```{r local_function_load}
# load function from local files
source(here::here("source", "DEG_functions.R"))
source(here::here("source", "GSEA_analysis.R"))
```

\newpage

# 1. Read the count data
In this section, we will read the clean count data from the synaptosomes_bulkRNA folder.  We will read the data and merge them into a single table. 

```{r load_bulkRNA_data}
input_count <- read.csv(here::here("data", "bulkRNA",
                                      "bulkRNA_counts_cleaned.csv"))
counts <-  as.data.frame(input_count) %>% 
  column_to_rownames(var = "gene")
colnames(counts) <- gsub("_", "-", colnames(counts))


# raw sample list
sample_list_raw <- read.csv(here::here("data", "bulkRNA",
                                      "sample_info_FXS.csv")) %>%
                    mutate(condition = paste0(Diagosis, "_", Treatment),
                           sample = gsub("_", "-", Sample.name))


# Ensure the column names of counts exist in Sample.name
new_colnames <- sample_list_raw$Label[match(colnames(counts), sample_list_raw$sample )]

# Assign new column names
colnames(counts) <- new_colnames


# sort the columns by the colname
condition_list <- data.frame(
  group =sample_list_raw$condition
)

row.names(condition_list) <- sample_list_raw$Label

counts<- counts[, rownames(condition_list)]

gene_name_mapping<- readRDS(here::here("data","ref" ,"gene_name_mapping.rds"))
```



```{r DESeq2_analysis}
# Init the result folder structure for the result
result_folder_all = './results/04-BioDomain'
result_folder = result_folder_all

```


\newpage

# 2. Visualization for reuslt

## (1) Sample information - PCA plot

```{r Sample_PCA, fig.width=8, fig.height=6}
figure_folder = result_folder
# do PCA for counts data
dds_obj <- DESeqDataSetFromMatrix(countData = counts,
                                  colData = condition_list,
                                  design = ~ group)
vsd.obj <- varianceStabilizingTransformation(dds_obj, blind = TRUE)
pcaData <- plotPCA(vsd.obj,  intgroup = c("group"), returnData = T)
percentVar <- round(100 * attr(pcaData, "percentVar"))


p <-ggplot(pcaData, aes(PC1, PC2, color=group)) +
  geom_point(size=3) +
  labs(x = paste0("PC1: ",percentVar[1],"% variance"),
       y = paste0("PC2: ",percentVar[2],"% variance"),
  ) +
  stat_ellipse(level = 0.95)+
  theme_bw() +
  # theme_classic()+
  theme(text = element_text(family = "Arial", colour = "black")) +
  # scale_color_manual(values = assigned_colors) +
  ggrepel::geom_text_repel(aes(label = name), color = "black")

print(p)
# ggsave("./results/01-Sample_info/01_sample_PCA_plot.pdf", p,width = 8, height = 6, units = "in", dpi = 300)
# ggsave("./results/01-Sample_info/01_sample_PCA_plot.png", p,width = 8, height = 6, units = "in", dpi = 300)
#   
```

## (2) Sample information - Distance heatmap

```{r Sample_dis_Heatmap, fig.width=8, fig.height=6}
 # Now apply variance stabilizing transformation
 sampleDists <- dist(t(assay(vsd.obj)))
 sampleDistMatrix <- as.matrix( sampleDists )
 rownames(sampleDistMatrix) <- paste( vsd.obj$group )
 colors <- colorRampPalette( rev(RColorBrewer::brewer.pal(9, "Blues")) )(255)
 p <- pheatmap::pheatmap(sampleDistMatrix,
                         clustering_distance_rows = sampleDists,
                         clustering_distance_cols = sampleDists,
                         col = colors) 
print(p)

#  ggsave("./results/01-Sample_info/02_sample_distance_heatmap.pdf", p,width = 8, height = 6, units = "in", dpi = 300)
# ggsave("./results/01-Sample_info/02_sample_distance_heatmap.png",
#        p, width = 8, height = 6, units = "in", dpi = 300)

```

\newpage

# 3. GSEA analysis and BioDomain analysis

### (1) GSEA analysis for AD vs CTRL

```{r GSEA_analysis_AD}
# result_folder= "./results/04-BioDomain"
# result_df <- perform_gsea_analysis(counts, condition_list, 
#                                    ref_group = "CTRL_Veh_0", 
#                                    compare_group = "AD_Veh_0")
# write.csv(result_df, file.path(result_folder, "GSEA_AD_vs_CTRL.csv"))
```

### (2) GSEA analysis for AD_Drug vs CTRL

```{r GSEA_analysis_AD_Drug}
# result_df <- perform_gsea_analysis(counts, condition_list, 
#                                    ref_group = "CTRL_Veh_0", 
#                                    compare_group = "AD_Veh_10")
# write.csv(result_df, file.path(result_folder, "GSEA_AD_Drug_vs_CTRL.csv"))
```


### (3) GSEA analysis for CTRL_Drug vs CTRL

```{r GSEA_analysis_CTRL_Drug}
# result_df <- perform_gsea_analysis(counts, condition_list, 
#                                    ref_group = "CTRL_Veh_0", 
#                                    compare_group = "CTRL_Veh_10")
# write.csv(result_df, file.path(result_folder, "GSEA_CTRL_Drug_vs_CTRL.csv"))

```


\newpage

# 4. BioDomain analysis

```{r}
ref_BioDomain <- read.csv(here::here("data", "ref", "BioDomain_Reference_2025_04_01.csv"))

result_AD <- read.csv(file.path(result_folder, "GSEA_AD_vs_CTRL.csv"), row.names = 1)
result_AD <- result_AD %>% mutate(AD_NES = NES, AD_pval = p.adjust) %>% 
  dplyr::select(ID,Description, AD_NES, AD_pval)


result_AD_Drug <- read.csv(file.path(result_folder, "GSEA_AD_Drug_vs_CTRL.csv"), row.names = 1)
result_AD_Drug <- result_AD_Drug %>% mutate(AD_Drug_NES = NES, AD_Drug_pval = p.adjust) %>% 
  dplyr::select(ID,Description, AD_Drug_NES, AD_Drug_pval)

result_CTRL_Drug <- read.csv(file.path(result_folder, "GSEA_CTRL_Drug_vs_CTRL.csv"), row.names = 1)
result_CTRL_Drug <- result_CTRL_Drug %>% mutate(CTRL_Drug_NES = NES, CTRL_Drug_pval = p.adjust) %>% 
  dplyr::select(ID,Description, CTRL_Drug_NES, CTRL_Drug_pval)

```



# 5. Load the GSVA analysis

```{r make_reference_for_GSVA}
# 
# # Install if not already
# if (!requireNamespace("msigdbr", quietly = TRUE)) {
#   install.packages("msigdbr")
# }
# library(msigdbr)
# 
# # Load all human GO sets (C5 category)
# msig_go <- msigdbr(species = "Homo sapiens", category = "C5")
# 
# # Extract and keep unique pairs of gene set name and GO ID
# go_pairs <- unique(msig_go[, c("gs_cat","gs_subcat","gs_name", "gs_exact_source")])
# 
# 
# # save the misi_go to data/ref folder with here::
# write.csv(go_pairs, here::here("data","ref","GO_reference_2025_04_02.csv"))

```





```{r}
# load the GSVA analysis result
result_GSVA <- read.csv(here::here("analysis", "03-Mutil_group_GSVA", "results","02-GSVA","01_GSVA_matrix.csv"), row.names = 1)

# group level average GSVA score
group_list <- c("AD_Veh_0", "AD_Veh_10", "CTRL_Veh_0", "CTRL_Veh_10")
condition_list_label <- condition_list %>% filter(group %in% group_list)

result_GSVA <- result_GSVA [, rownames(condition_list_label)]


result_GSVA_group <- tibble()
row.names(result_GSVA) <- row.names(result_GSVA)





library(dplyr)
library(tibble)

# Initialize an empty tibble to store group averages
result_GSVA_group <- matrix(NA, 
                            nrow = nrow(result_GSVA), 
                            ncol = length(group_list),
                            dimnames = list(rownames(result_GSVA), group_list)) 


# Loop through each group and calculate the average GSVA score
for (grp in group_list) {
  # Get sample names for the current group
  sample_names <- condition_list_label %>%
    filter(group == grp) %>%
    # trun rownames to sample name
    rownames()
  
  # Subset the GSVA matrix for the samples in this group
  group_data <- result_GSVA[, sample_names, drop = FALSE]
  
  # Compute the row means (i.e., average GSVA score per gene/pathway)
  avg_scores <- rowMeans(group_data, na.rm = TRUE)
  
  # Add to the result table
   result_GSVA_group[, grp] <-  avg_scores
}




# # Match based on the GO term text
# result_GSVA_group <-as.data.frame(result_GSVA_group)
# result_GSVA_group$ID <- rownames(result_GSVA_group)
# result_GSVA_group$ID <- gsub("_", " ", result_GSVA_group$ID)
# 
# # turn id to be lower case
# result_GSVA_group$ID <- tolower(result_GSVA_group$ID)
# result_GSVA_group$ID <- gsub("gobp ", "", result_GSVA_group$ID)
# result_GSVA_group$ID <- gsub("gocc ", "", result_GSVA_group$ID)
# 
# ref_BioDomain <- read.csv(here::here("data", "ref", "BioDomain_Reference_2025_04_01.csv"))
# ref_BioDomain$ID <- ref_BioDomain$GOterm_Name
# ref_BioDomain$ID <- tolower(ref_BioDomain$ID)
# 
# df_BioDomain <- left_join(ref_BioDomain,result_GSVA_group, by =  "ID")


# result_GSVA_group <- merge(result_GSVA_group, go_ref, by.x = "row.names", by.y = "row.names", all.x = TRUE)
# go_ref <- read.csv(here::here("data","ref","GO_reference_2025_04_02.csv"), row.names = 1)
# rownames(go_ref) <- go_ref$gs_name 



# merge the BioDomain with the GSVA result
result_GSVA_group_df <- as.data.frame(result_GSVA_group)
# result_GSVA_group_df <- left_join(result_GSVA_group_df ,ref_pairs , by.x =  "ID", by.y = "gs_name", all.x = TRUE)



go_ref <- read.csv(here::here("data","ref","GO_reference_2025_04_02.csv"), row.names = 1)
ref_BioDomain <- read.csv(here::here("data", "ref", "BioDomain_Reference_2025_04_01.csv"))


# check if the GO_ID is unique in the ref_BioDomain
length(unique(ref_BioDomain$GO_ID)) == nrow(ref_BioDomain)


ref_BioDomain_go <- merge(
  go_ref,
  ref_BioDomain,
  by.x = "gs_exact_source",
  by.y = "GO_ID",
  all.x = TRUE
)

dim(ref_BioDomain_go)
head(ref_BioDomain_go)

# chcek the NA in the Biodomain
sum(is.na(ref_BioDomain_go$Biodomain))

# fill the NA with "Unknown"
ref_BioDomain_go$Biodomain[is.na(ref_BioDomain_go$Biodomain)] <- "Unknown"
ref_pairs <- ref_BioDomain_go[, c("gs_name", "Biodomain")]


result_GSVA_group_df$ID <- rownames(result_GSVA_group_df)
result_GSVA_group_df <- left_join(
  result_GSVA_group_df,
  ref_pairs,
  by = c("ID" = "gs_name")
)
# fill the NA with "Unknown"
result_GSVA_group_df$Biodomain[is.na(result_GSVA_group_df$Biodomain)] <- "Unknown"

```

## (1) Ridge plot for GSVA score
```{r BioDomain_Ridge,fig.width=24, fig.height=20}

library(tidyverse)
library(ggridges)
library(ggdist)

# Reshape data to long format
df_long <- result_GSVA_group_df %>%
  pivot_longer(
    cols = c( CTRL_Veh_0, CTRL_Veh_10,AD_Veh_0, AD_Veh_10),
    names_to = "Group",
    values_to = "GSVA_Score"
  ) %>%
  filter(!is.na(GSVA_Score)) %>%
  filter(Biodomain != "Unknown")




df_long$group <- fct_recode(df_long$Group,
  "CTRL_Veh" = "CTRL_Veh_0",
  "CTRL_10" = "CTRL_Veh_10",
  "AD_Veh" = "AD_Veh_0",
  "AD_10" = "AD_Veh_10"
)


df_long$group <- factor(df_long$group, levels = c( "AD_10","AD_Veh","CTRL_10","CTRL_Veh" ))

# Ridge plot
p <- ggplot(df_long, aes(x = GSVA_Score, y = group, fill = group)) +
  geom_density_ridges(alpha = 0.7, scale = 1.2, rel_min_height = 0.01) +
  facet_wrap(~ Biodomain, scales = "free_y") +
  labs(title = "Ridge Plot of GSVA Scores by Biodomain and Group",
       x = "GSVA Score", y = "Group") +
  theme_bw() +
  theme(legend.position = "none",
        strip.text = element_text(size = 12),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10))

print(p)


ggsave(file.path(result_folder, "GSVA_Ridge_plot_v2.png"), p, width = 24, height = 20, units = "in", dpi = 300)
ggsave(file.path(result_folder, "GSVA_Ridge_plot_v2.pdf"), p, width = 24, height = 20, units = "in", dpi = 300)

```
\newpage
## (2) Raincloud plot for GSVA score

```{r BioDomain_Cloud, fig.width=24, fig.height=20}


## (2) Raincloud plot for GSVA score (suppressing verbose output)
suppressWarnings({
  p <- ggplot(df_long, aes(x = GSVA_Score, y = group, fill = group)) +
    stat_halfeye(
      adjust = 0.6,
      justification = -0.3,
      .width = 0,
      point_colour = NA
    ) +
    geom_boxplot(
      width = 0.15,
      outlier.shape = NA,
      alpha = 0.4
    ) +
    geom_jitter(
      aes(color = group),
      width = 0.001,
      size = 0.8,
      alpha = 0.4
    ) +
    facet_wrap(~ Biodomain, scales = "free_y") +
    labs(
      title = "GSVA Distribution per Biodomain (Raincloud Plot)",
      x = "GSVA Score", y = NULL
    ) +
    theme_bw() +
    theme(
      legend.position = "none",
      strip.text = element_text(size = 13),
      axis.text.y = element_text(size = 11),
      axis.text.x = element_text(size = 10)
    )
})



print(p)

ggsave(file.path(result_folder, "GSVA_cloud_plot_v2.png"),  width = 24, height = 20, units = "in", dpi = 300)
ggsave(file.path(result_folder, "GSVA_cloud_plot_v2.pdf"),  width = 24, height = 20, units = "in", dpi = 300)


```

\newpage
## (3) Rug and density plot for GSVA score

```{r BioDomain_Rug, fig.width=24, fig.height=20}
## (2) Ridge + Rug plot for GSVA score

library(ggridges)

p <- ggplot(df_long, aes(x = GSVA_Score, y = group, fill = group)) +
  geom_density_ridges(scale = 1.2, alpha = 0.6, color = "white") +
  geom_rug(aes(color = group), sides = "b", alpha = 0.4, size = 0.5) +
  facet_wrap(~ Biodomain, scales = "free_y") +
  labs(
    title = "GSVA Distribution per Biodomain (Ridge + Rug Plot)",
    x = "GSVA Score", y = NULL
  ) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 13),
    axis.text.y = element_text(size = 11),
    axis.text.x = element_text(size = 10)
  )

print(p)

ggsave(file.path(result_folder, "GSVA_ridge_rug_plot_v2.png"), width = 24, height = 20, units = "in", dpi = 300)
ggsave(file.path(result_folder, "GSVA_ridge_rug_plot_v2.pdf"), width = 24, height = 20, units = "in", dpi = 300)



```
\newpage
# Session information
```{r}
sessionInfo()
```

