---
title: "GSVA for mutil Group"
author: "Ximing Ran"
date: "2025-03-26"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: false
  html_document:
    # code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js

---

```{r setup, include=FALSE}
# load libraries
library(tidyverse)
library(knitr)
library(dplyr)
library(car)
library(lme4)
library(lmerTest)


set.seed(2024)

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.path = "./results/03-Mix-effect/Analysis_figure/"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
knitr::kable(head(mtcars[, 1:4]), "simple")
```

```{r}
library(tibble)
library(tidyr)
library(dplyr)
library(rtracklayer)
library(cowplot)
```



```{r local_function_load}
# load function from local files
source(here::here("source", "DEG_functions.R"))
```

\newpage

# 1. Read the count data
In this section, we will read the clean count data from the synaptosomes_bulkRNA folder.  We will read the data and merge them into a single table. 

```{r load_bulkRNA_data}
input_count <- read.csv(here::here("data", "bulkRNA",
                                      "bulkRNA_counts_cleaned.csv"))
counts <-  as.data.frame(input_count) %>% column_to_rownames(var = "gene")
colnames(counts) <- gsub("_", "-", colnames(counts))

# if the colname is start with X, remove it
colnames(counts) <- gsub("^X", "", colnames(counts))

# raw sample list
sample_list_raw <- read.csv(here::here("data", "bulkRNA",
                                      "sample_info_FXS.csv")) %>%
                    mutate(condition = Label,
                           sample = gsub("_", "-", Sample_ID)) %>%
                    filter(Diagosis != "ISO")


# Ensure the column names of counts exist in Sample.name
new_colnames <- sample_list_raw$sample[match(colnames(counts), sample_list_raw$sample )]

# Assign new column names
colnames(counts) <- new_colnames


# sort the columns by the colname
condition_list <- data.frame(
  group =sample_list_raw$condition
)

row.names(condition_list) <- sample_list_raw$sample

counts<- counts[, rownames(condition_list)]

gene_name_mapping<- readRDS(here::here("data","ref" ,"gene_name_mapping.rds"))




```


```{r DESeq2_analysis}
# Init the result folder structure for the result
result_folder_all = './results'
result_folder = result_folder_all

```


# 2. Visualization for Result

## (1) Sample Information - PCA Plot

![PCA Plot with individual label](./results/01-Sample_info/01_sample_PCA_plot.pdf)

\newpage

![PCA Plot with group label](./results/01-Sample_info/01_sample_PCA_plot_group.pdf)

\newpage

## (2) Sample Information - Distance Heatmap

![Sample Information Distance Heatmap](./results/01-Sample_info/02_sample_distance_heatmap.pdf)

\newpage


# 3. GSVA analysis with Mix effect model.
## (1) Main Method: Mix-Effects Model with no Interaction

We model **GSVA scores** using a **mixed-effects model**, incorporating **fixed effects** for disease status (FXS) and drug treatment, and a **random intercept** for individual-level variability:

\[
Y_{ij} = \beta_0 + \beta_1 X_{\text{FXS}} + \beta_2 X_{\text{Drug\_BAY}} + \beta_3 X_{\text{Drug\_BPN}} + \beta_4 X_{\text{Drug\_BP}} + b_i + \varepsilon_{ij}
\]

### **Model Components**
- \(Y_{ij}\): GSVA score for individual \(i\) under condition \(j\).

- **Fixed Effects:**
  - \(X_{\text{FXS}}\): Disease indicator (1 for **FXS**, 0 for **CTRL**).
  - \(X_{\text{Drug\_BAY}}, X_{\text{Drug\_BPN}}, X_{\text{Drug\_BP}}\): Drug treatment indicators (vehicle is reference).
  
- **Random intercept** 
  - \(b_i \sim \mathcal{N}(0, \tau^2)\) for individual \(i\), capturing baseline variability.
  - \(\varepsilon_{ij} \sim \mathcal{N}(0, \sigma^2)\): Residual error term.

This model allows us to estimate the effects of disease and drug treatments on GSVA scores while controlling for repeated measures within individuals.



\newpage
```{r GSVA_analysis, fig.width=6, fig.height=6}

#
# # The following code is used to generate the GSVA matrix , only need to run once
# gmxFile <- here::here("data", "ref", "c5.go.v2023.1.Hs.symbols.gmt")
# go_list <- getGmt(gmxFile)
# 
# geneset <- go_list
# dat <- as.matrix(counts)
# 
# gsvapar <- gsvaParam(dat, geneset, maxDiff=TRUE)
# gsva_es <- gsva(gsvapar)
# gsva_matrix <- as.data.frame(gsva_es)
# 
# # save the result
# write.csv(gsva_matrix, "./results/02-GSVA/01_GSVA_matrix.csv")
# 

# plot the heatmap for the GSVA result
# pathway_list <- read.csv(here::here("data", "ref", "focus-pathway_2024_10_03.csv"))
pathway_list <- read.csv(here::here("data", "ref", "focus-pathway_2025_03_26.csv"))

box_plot_folder<- file.path(result_folder,"04-GSVA","Boxplot")
# create the folder
dir.create(box_plot_folder, showWarnings = FALSE)

gsva_matrix <- read.csv("./results/02-GSVA/01_GSVA_matrix.csv", row.names = 1)
colnames(gsva_matrix) <- sub("^X", "", colnames(gsva_matrix))
colnames(gsva_matrix) <- gsub("\\.", "-", colnames(gsva_matrix))
condition_list_label <- condition_list 
condition_list_label$group <- factor(
  condition_list_label$group,
  # levels = c("CTRL_Veh_0", "CTRL_Veh_10",  "CTRL_Veh_100",
  #            "AD_Veh_0", "AD_Veh_10", "AD_Veh_100")
  levels = c("CTRL_VEH", "CTRL_BAY", "CTRL_BPN","CTRL_BP",
             "FXS_VEH",  "FXS_BAY",  "FXS_BPN", "FXS_BP")
)



sample_df_clean <- sample_list_raw %>% dplyr::select(Sample_ID,Individual, Diagosis, Treatment) %>% 
  mutate(X_FXS = ifelse(Diagosis == "FXS", 1, 0),
         X_Drug_BAY = ifelse(Treatment == "BAY", 1, 0),
         X_Drug_BPN = ifelse(Treatment == "BPN", 1, 0),
         X_Drug_BP = ifelse(Treatment == "BP", 1, 0))



pathway_name <- "GOBP_CELL_CYCLE_CHECKPOINT_SIGNALING"

  
# Step 1: Prepare the data for the model
gsva_df <- as.data.frame(t(gsva_matrix[pathway_name,])) %>%
    mutate(Sample_ID = rownames(.)) 
colnames(gsva_df) <-c("Y", "Sample_ID")

# Step 2: Merge by Sample_ID
df_pathway <- left_join(sample_df_clean,gsva_df,  by = "Sample_ID")


rg_model <- lmer(Y ~ X_FXS + X_Drug_BAY + X_Drug_BPN + X_Drug_BP + (1 | Individual), data = df_pathway)


summary_model <- summary(rg_model)
print(summary_model)


```
\newpage

## (2) Main Method: Mix-Effects Model with Interaction

We model **GSVA scores** using a **mixed-effects model**, incorporating **fixed effects** for disease status (FXS), drug treatments, and their **interactions**, along with a **random intercept** for individual-level variability:

\[
\begin{aligned}
Y_{ij} =\ & \beta_0 + \beta_1 X_{\text{FXS}} + \beta_2 X_{\text{Drug\_BAY}} + \beta_3 X_{\text{Drug\_BPN}} + \beta_4 X_{\text{Drug\_BP}} \\
          & + \beta_5 (X_{\text{FXS}} \times X_{\text{Drug\_BAY}}) + \beta_6 (X_{\text{FXS}} \times X_{\text{Drug\_BPN}}) + \beta_7 (X_{\text{FXS}} \times X_{\text{Drug\_BP}}) \\
          & +\ b_i + \varepsilon_{ij}
\end{aligned}
\]


### **Model Components**
- \(Y_{ij}\): GSVA score for individual \(i\) under condition \(j\).

- **Fixed Effects:**
  - \(X_{\text{FXS}}\): Disease indicator (1 for **FXS**, 0 for **CTRL**).
  - \(X_{\text{Drug\_BAY}}, X_{\text{Drug\_BPN}}, X_{\text{Drug\_BP}}\): Drug treatment indicators (vehicle is the reference).
  - **Interaction Terms:**
    - \(X_{\text{FXS}} \times X_{\text{Drug\_BAY}}\): Interaction between FXS and BAY treatment.
    - \(X_{\text{FXS}} \times X_{\text{Drug\_BPN}}\): Interaction between FXS and BPN treatment.
    - \(X_{\text{FXS}} \times X_{\text{Drug\_BP}}\): Interaction between FXS and BP treatment.

- **Random Intercept:**
  - \(b_i \sim \mathcal{N}(0, \tau^2)\): Individual-specific random intercept capturing baseline variability.
  
- **Error Term:**
  - \(\varepsilon_{ij} \sim \mathcal{N}(0, \sigma^2)\): Residual error term.

This model allows us to estimate not only the **main effects** of disease and drug treatments but also **how drug responses differ between FXS and CTRL individuals**. Interaction terms capture whether the drug effect is **modified by disease status**, which is essential for understanding differential pathway responses.

\newpage
```{r}

rg_model <- lmer(Y ~ X_FXS + X_Drug_BAY + X_Drug_BPN + X_Drug_BP +
                   X_FXS*X_Drug_BAY + X_FXS*X_Drug_BPN + X_FXS*X_Drug_BP +
                   (1 | Individual), data = df_pathway)


summary_model <- summary(rg_model)
print(summary_model)

```


\newpage
### Hypothesis Testing: Comparing differece with FXS after Drug and CTRL 

### **Combined Effect Hypothesis Testing**

To interpret the effect of each drug within the **FXS background**, we tested whether the **combined effect** of disease and drug treatment is significantly different from CTRL. Specifically, we evaluated the null hypothesis:

\[
H_0: \beta_{\text{FXS}} + \beta_{\text{Drug}} + \beta_{\text{FXS:Drug}} = 0
\]

This tests whether the **net drug effect in the FXS group** is equivalent to the CTRL baseline (i.e., no rescue effect).


```{r}
# Load required package
linearHypothesis(
  rg_model,hypothesis.matrix = "X_FXS +X_Drug_BAY + X_FXS:X_Drug_BAY = 0"
)

# Define and test the hypothesis: beta_FXS + beta_FXS:Drug_BPN = 0
linearHypothesis(
  rg_model,hypothesis.matrix = "X_FXS +X_Drug_BPN+ X_FXS:X_Drug_BPN = 0"
)

linearHypothesis(
  rg_model,hypothesis.matrix = "X_FXS + X_Drug_BP + X_FXS:X_Drug_BP = 0"
)

```


```{r}
# library(dplyr)
# library(lme4)
# library(car)
# 
# pathway_list_all <- rownames(gsva_matrix)
# 
# # Step 0: Clean metadata
# sample_df_clean <- sample_list_raw %>%
#   dplyr::select(Sample_ID, Individual, Diagosis, Treatment) %>%
#   mutate(
#     X_FXS = ifelse(Diagosis == "FXS", 1, 0),
#     X_Drug_BAY = ifelse(Treatment == "BAY", 1, 0),
#     X_Drug_BPN = ifelse(Treatment == "BPN", 1, 0),
#     X_Drug_BP  = ifelse(Treatment == "BP", 1, 0)
#   )
# 
# # Step 1: Initialize result storage
# result_df <- data.frame(
#   pathway = character(),
#   beta_FXS = numeric(),
#   beta_Drug_BAY = numeric(),
#   beta_Drug_BPN = numeric(),
#   beta_Drug_BP = numeric(),
#   beta_FXS_Drug_BAY = numeric(),
#   beta_FXS_Drug_BPN = numeric(),
#   beta_FXS_Drug_BP = numeric(),
#   p_FXS = numeric(),
#   p_Drug_BAY = numeric(),
#   p_Drug_BPN = numeric(),
#   p_Drug_BP = numeric(),
#   p_FXS_Drug_BAY = numeric(),
#   p_FXS_Drug_BPN = numeric(),
#   p_FXS_Drug_BP = numeric(),
#   combined_p_BAY = numeric(),
#   combined_p_BPN = numeric(),
#   combined_p_BP = numeric(),
#   stringsAsFactors = FALSE
# )
# 
# # Step 2: Loop through all pathways
# for (i in 1:length(pathway_list_all)) {
# # for (i in 1:5) {
#   
#   # send notification for every 100 pathways
#   if (i %% 100 == 0) {
#     cat("Processing pathway", i, "of",length(pathway_list_all), "\n")
#   }
#   
#   pathway_name <- pathway_list_all[i]
#   
#   # cat("Processing:", pathway_name, "\n")
#   
#   # Extract GSVA scores for this pathway
#   gsva_df <- as.data.frame(t(gsva_matrix[pathway_name, ])) %>%
#     mutate(Sample_ID = rownames(.))
#   colnames(gsva_df) <- c("Y", "Sample_ID")
#   
#   # Merge with sample metadata
#   df_pathway <- left_join(sample_df_clean, gsva_df, by = "Sample_ID")
#   
#   # Make sure random effect variable is factor
#   df_pathway$Individual <- as.factor(df_pathway$Individual)
#   
#   # Fit mixed-effects model
#   model <- tryCatch({
#     lmer(Y ~ X_FXS + X_Drug_BAY + X_Drug_BPN + X_Drug_BP +
#            X_FXS * X_Drug_BAY + X_FXS * X_Drug_BPN + X_FXS * X_Drug_BP +
#            (1 | Individual), data = df_pathway)
#   }, error = function(e) return(NULL))
#   
#   if (is.null(model)) next  # Skip if model fails
#   
#   coefs <- coef(summary(model))
#   
#   # Safe extraction
#   get_coef <- function(name) if (name %in% rownames(coefs)) coefs[name, "Estimate"] else NA
#   get_pval <- function(name) if (name %in% rownames(coefs)) coefs[name, "Pr(>|t|)"] else NA
#   
#   # Combined effects using linearHypothesis
#   p_comb_bay <- tryCatch({
#     linearHypothesis(model, "X_FXS + X_Drug_BAY + X_FXS:X_Drug_BAY = 0")$`Pr(>Chisq)`[2]
#   }, error = function(e) NA)
#   
#   p_comb_bpn <- tryCatch({
#     linearHypothesis(model, "X_FXS + X_Drug_BPN + X_FXS:X_Drug_BPN = 0")$`Pr(>Chisq)`[2]
#   }, error = function(e) NA)
#   
#   p_comb_bp <- tryCatch({
#     linearHypothesis(model, "X_FXS + X_Drug_BP + X_FXS:X_Drug_BP = 0")$`Pr(>Chisq)`[2]
#   }, error = function(e) NA)
# 
#   
#   # Store results
#   result_df <- rbind(result_df, data.frame(
#     pathway = pathway_name,
#     beta_FXS = get_coef("X_FXS"),
#     beta_Drug_BAY = get_coef("X_Drug_BAY"),
#     beta_Drug_BPN = get_coef("X_Drug_BPN"),
#     beta_Drug_BP = get_coef("X_Drug_BP"),
#     beta_FXS_Drug_BAY = get_coef("X_FXS:X_Drug_BAY"),
#     beta_FXS_Drug_BPN = get_coef("X_FXS:X_Drug_BPN"),
#     beta_FXS_Drug_BP = get_coef("X_FXS:X_Drug_BP"),
#     p_FXS = get_pval("X_FXS"),
#     p_Drug_BAY = get_pval("X_Drug_BAY"),
#     p_Drug_BPN = get_pval("X_Drug_BPN"),
#     p_Drug_BP = get_pval("X_Drug_BP"),
#     p_FXS_Drug_BAY = get_pval("X_FXS:X_Drug_BAY"),
#     p_FXS_Drug_BPN = get_pval("X_FXS:X_Drug_BPN"),
#     p_FXS_Drug_BP = get_pval("X_FXS:X_Drug_BP"),
#     combined_p_BAY = p_comb_bay,
#     combined_p_BPN = p_comb_bpn,
#     combined_p_BP = p_comb_bp,
#     stringsAsFactors = FALSE
#   ))
# }
# 
# # View or save results
# # View(result_df)
# write.csv(result_df, "./results/03-Mix-effect/mixed_model_results_all.csv", row.names = FALSE)


```

\newpage

## (3) Pathway Analysis: Full Rescue, Partial Rescue, and Side Effects

### a. BAY Treatment

```{r BAY, fig.width=12, fig.height=6}
result_df <- read.csv("./results/03-Mix-effect/mixed_model_results_all.csv")

# Step0 : Select pathaways with signficant FXS effect
result_df <- result_df %>%
  filter(p_FXS < 0.05) %>%
  arrange(p_FXS)


#----------------------------------------#
#             Drug :   BAY               #
#----------------------------------------#

# Step1 : Select pathaways with signficant Drug reverse effect
# Action : filter the X_FXS:X_Drug_BAY is signficant and is on the reverse direction with X_FXS

result_df_BAY <- result_df %>%
  filter(p_FXS_Drug_BAY < 0.05 & beta_FXS * beta_FXS_Drug_BAY < 0) %>%
  arrange(p_FXS_Drug_BAY)

dim(result_df_BAY)


# Step2 : Define the degree of rescue
# Action:
# We define two rescue categories based on statistical comparisons:
# - Full Rescue: If the post-treatment average is significantly different from the pre-treatment average
#                and NOT significantly different from the CTRL group, we classify it as a "full rescue".
# - Partial Rescue: If the post-treatment average is significantly different from the pre-treatment average
#                   but still significantly different from the CTRL group, we classify it as a "partial rescue".


result_df_BAY <- result_df_BAY  %>%
  mutate(rescue = case_when(
    p_FXS_Drug_BAY < 0.05 & combined_p_BAY >= 0.05 ~ "Full Rescue",
    p_FXS_Drug_BAY < 0.05 & combined_p_BAY < 0.05 ~ "Partial Rescue",
    TRUE ~ "No Rescue"
  ))


# Step3 : Consider Side effect
# Action: If the CTRL After treatment is significantly different from the CTRL Before treatment, we classify it as a "side effect".

result_df_BAY <- result_df_BAY  %>%
  mutate(side_effect = case_when(
    p_Drug_BAY < 0.05 ~ "Side Effect",
    TRUE ~ "No Side Effect"
  ))
result_df_BAY <- result_df_BAY %>% filter(side_effect == "No Side Effect") %>% dplyr::select(pathway,rescue, side_effect)
write.csv(result_df_BAY, "./results/03-Mix-effect/mixed_model_results_BAY.csv", row.names = FALSE)

condition_list_label_BAY <- condition_list_label %>% filter(group %in% c("CTRL_VEH", "CTRL_BAY", "FXS_VEH", "FXS_BAY"))

# # Only need to run once
# for (i in 1:nrow(result_df_BAY)) {
#   if ( i %% 10 == 0) {
#     cat("Processing pathway", i, "of", nrow(result_df_BAY), "\n")
#   }
#   pathway_name <- result_df_BAY$pathway[i]
# 
#   p<-plot_gsva_boxplot_mutil(gsva_matrix,
#                       condition_list_label = condition_list_label,
#                       pathway_name =  pathway_name,
#                       figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_BAY_all_DRUG"),
#                       file_name = paste0("GSVA_", pathway_name, "_DRUG"),
#                       fig.height = 6, fig.width = 6)
# 
# 
#   p<-plot_gsva_boxplot_mutil(gsva_matrix,
#                       condition_list_label = condition_list_label_BAY,
#                       pathway_name =  pathway_name,
#                       figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot-BAY"),
#                       file_name = paste0("GSVA_", pathway_name, "_BAY"),
#                       fig.height = 6, fig.width = 4)
# }


# Only need to run once
for (i in 1:2) {
  if ( i %% 10 == 0) {
    cat("Processing pathway", i, "of", nrow(result_df_BAY), "\n")
  }
  pathway_name <- result_df_BAY$pathway[i]

  p1<-plot_gsva_boxplot_mutil(gsva_matrix,
                      condition_list_label = condition_list_label,
                      pathway_name =  pathway_name,
                      figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_BAY_all_DRUG"),
                      file_name = paste0("GSVA_", pathway_name, "_DRUG"),
                      fig.height = 6, fig.width = 6)



  p2<-plot_gsva_boxplot_mutil(gsva_matrix,
                      condition_list_label = condition_list_label_BAY,
                      pathway_name =  pathway_name,
                      figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot-BAY"),
                      file_name = paste0("GSVA_", pathway_name, "_BAY"),
                      fig.height = 6, fig.width = 4)


  
p_comp <- plot_grid(p1, p2, labels = c("A", "B"))
print(p_comp)

}


```


\newpage
### b. BPN Treatment

```{r BPN, fig.width=12, fig.height=6}

#----------------------------------------#
#             Drug :   BPN               #
#----------------------------------------#

# Step1 : Select pathaways with signficant Drug reverse effect
# Action : filter the X_FXS:X_Drug_BPN is signficant and is on the reverse direction with X_FXS

result_df_BPN <- result_df %>%
  filter(p_FXS_Drug_BPN < 0.05 & beta_FXS * beta_FXS_Drug_BPN < 0) %>%
  arrange(p_FXS_Drug_BPN)

# Step2 : Define the degree of rescue
# Action:
# We define two rescue categories based on statistical comparisons:
# - Full Rescue: If the post-treatment average is significantly different from the pre-treatment average
#                and NOT significantly different from the CTRL group, we classify it as a "full rescue".
# - Partial Rescue: If the post-treatment average is significantly different from the pre-treatment average
#                   but still significantly different from the CTRL group, we classify it as a "partial rescue".

result_df_BPN <- result_df_BPN  %>%
  mutate(rescue = case_when(
    p_FXS_Drug_BAY < 0.05 & combined_p_BPN >= 0.05 ~ "Full Rescue",
    p_FXS_Drug_BAY < 0.05 & combined_p_BPN < 0.05 ~ "Partial Rescue",
    TRUE ~ "No Rescue"
  ))

# Step3 : Consider Side effect
# Action: If the CTRL After treatment is significantly different from the CTRL Before treatment, we classify it as a "side effect".

result_df_BPN <- result_df_BPN  %>%
  mutate(side_effect = case_when(
    p_Drug_BPN < 0.05 ~ "Side Effect",
    TRUE ~ "No Side Effect"
  ))

result_df_BPN <- result_df_BPN %>% filter(side_effect == "No Side Effect") %>% dplyr::select(pathway,rescue, side_effect)
write.csv(result_df_BPN, "./results/03-Mix-effect/mixed_model_results_BPN.csv", row.names = FALSE)

condition_list_label_BPN <- condition_list_label %>% filter(group %in% c("CTRL_VEH", "CTRL_BPN", "FXS_VEH", "FXS_BPN"))

# for (i in 1:nrow(result_df_BPN)) {
#   if ( i %% 10 == 0) {
#     cat("Processing pathway", i, "of", nrow(result_df_BPN), "\n")
#   }
#   pathway_name <- result_df_BPN$pathway[i]
# 
#   plot_gsva_boxplot_mutil(gsva_matrix,
#                       condition_list_label = condition_list_label,
#                       pathway_name =  pathway_name,
#                       figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_BPN_all_DRUG"),
#                       file_name = paste0("GSVA_", pathway_name, "_DRUG"),
#                       fig.height = 6, fig.width = 6)
#   
#   
#   plot_gsva_boxplot_mutil(gsva_matrix,
#                       condition_list_label = condition_list_label_BPN,
#                       pathway_name =  pathway_name,
#                       figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot-BPN"),
#                       file_name = paste0("GSVA_", pathway_name, "_BPN"),
#                       fig.height = 6, fig.width = 4)
# }


for (i in 1:2 ){
  if ( i %% 10 == 0) {
    cat("Processing pathway", i, "of", nrow(result_df_BPN), "\n")
  }
  pathway_name <- result_df_BPN$pathway[i]

  p1<-plot_gsva_boxplot_mutil(gsva_matrix,
                      condition_list_label = condition_list_label,
                      pathway_name =  pathway_name,
                      figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_BPN_all_DRUG"),
                      file_name = paste0("GSVA_", pathway_name, "_DRUG"),
                      fig.height = 6, fig.width = 6)


  p2<-plot_gsva_boxplot_mutil(gsva_matrix,
                      condition_list_label = condition_list_label_BPN,
                      pathway_name =  pathway_name,
                      figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot-BPN"),
                      file_name = paste0("GSVA_", pathway_name, "_BPN"),
                      fig.height = 6, fig.width = 4)
p_comp <- plot_grid(p1, p2, labels = c("A", "B"))
print(p_comp)
}


```



\newpage
### c. BP Treatment

```{r BP, fig.width=12, fig.height=6}


#----------------------------------------#
#             Drug :   BP               #
#----------------------------------------#

# Step1 : Select pathaways with signficant Drug reverse effect
# Action : filter the X_FXS:X_Drug_BP is signficant and is on the reverse direction with X_FXS

result_df_BP <- result_df %>%
  filter(p_FXS_Drug_BP < 0.05 & beta_FXS * beta_FXS_Drug_BP < 0) %>%
  arrange(p_FXS_Drug_BP)

# Step2 : Define the degree of rescue
# Action:
# We define two rescue categories based on statistical comparisons:
# - Full Rescue: If the post-treatment average is significantly different from the pre-treatment average
#                and NOT significantly different from the CTRL group, we classify it as a "full rescue".
# - Partial Rescue: If the post-treatment average is significantly different from the pre-treatment average
#                   but still significantly different from the CTRL group, we classify it as a "partial rescue".

result_df_BP <- result_df_BP  %>%
  mutate(rescue = case_when(
    p_FXS_Drug_BAY < 0.05 & combined_p_BP >= 0.05 ~ "Full Rescue",
    p_FXS_Drug_BAY < 0.05 & combined_p_BP < 0.05 ~ "Partial Rescue",
    TRUE ~ "No Rescue"
  ))

# Step3 : Consider Side effect
# Action: If the CTRL After treatment is significantly different from the CTRL Before treatment, we classify it as a "side effect".

result_df_BP <- result_df_BP  %>%
  mutate(side_effect = case_when(
    p_Drug_BP < 0.05 ~ "Side Effect",
    TRUE ~ "No Side Effect"
  ))

result_df_BP <- result_df_BP %>% filter(side_effect == "No Side Effect") %>% dplyr::select(pathway,rescue, side_effect)
write.csv(result_df_BP, "./results/03-Mix-effect/mixed_model_results_BP.csv", row.names = FALSE)

condition_list_label_BP <- condition_list_label %>% filter(group %in% c("CTRL_VEH", "CTRL_BP", "FXS_VEH", "FXS_BP"))

# for (i in 1:nrow(result_df_BP)) {
#   if ( i %% 10 == 0) {
#     cat("Processing pathway", i, "of", nrow(result_df_BP), "\n")
#   }
#   pathway_name <- result_df_BP$pathway[i]
# 
#   plot_gsva_boxplot_mutil(gsva_matrix,
#                       condition_list_label = condition_list_label,
#                       pathway_name =  pathway_name,
#                       figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_BP_all_DRUG"),
#                       file_name = paste0("GSVA_", pathway_name, "_DRUG"),
#                       fig.height = 6, fig.width = 6)
# 
# 
#   plot_gsva_boxplot_mutil(gsva_matrix,
#                       condition_list_label = condition_list_label_BP,
#                       pathway_name =  pathway_name,
#                       figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot-BP"),
#                       file_name = paste0("GSVA_", pathway_name, "_BP"),
#                       fig.height = 6, fig.width = 4)
# }


for (i in 1:2) {
  if ( i %% 10 == 0) {
    cat("Processing pathway", i, "of", nrow(result_df_BP), "\n")
  }
  pathway_name <- result_df_BP$pathway[i]
  
  p1<-plot_gsva_boxplot_mutil(gsva_matrix,
                          condition_list_label = condition_list_label,
                          pathway_name =  pathway_name,
                          figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_BP_all_DRUG"),
                          file_name = paste0("GSVA_", pathway_name, "_DRUG"),
                          fig.height = 6, fig.width = 6)


  p2<-plot_gsva_boxplot_mutil(gsva_matrix,
                          condition_list_label = condition_list_label_BP,
                          pathway_name =  pathway_name,
                          figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot-BP"),
                          file_name = paste0("GSVA_", pathway_name, "_BP"),
                          fig.height = 6, fig.width = 4)
  
p_comp <- plot_grid(p1, p2, labels = c("A", "B"))
print(p_comp)
}


```

\newpage
### (4) Common Rescue Pathways

**All 3 drugs**. 




```{r All_3, fig.width=6, fig.height=6}
pathway_overlap_3 <- unique(result_df_BAY$pathway) %>% intersect(unique(result_df_BPN$pathway)) %>% intersect(unique(result_df_BP$pathway))



for (i in 1:length(pathway_overlap_3)) {
  if ( i %% 10 == 0) {
    cat("Processing pathway", i, "of", length(pathway_overlap_3), "\n")
  }
  pathway_name <- pathway_overlap_3[i]
  
  p<-plot_gsva_boxplot_mutil(gsva_matrix,
                          condition_list_label = condition_list_label,
                          pathway_name =  pathway_name,
                          figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_Overlap_3"),
                          file_name = paste0("GSVA_", pathway_name, "_DRUG"),
                          fig.height = 6, fig.width = 6)
  print(p)
  

}

write.csv(pathway_overlap_3, "./results/03-Mix-effect/pathway_overlap_3.csv", row.names = FALSE)


```

\newpage
#### Overlap 2 drugs

**BAY and BPN**


```{r BAY_BPN, fig.width=12, fig.height=6}

pathway_overlap_BAY_BPN <- unique(result_df_BAY$pathway) %>% intersect(unique(result_df_BPN$pathway))


condition_list_label_BAY_BPN <- condition_list_label %>% filter(group %in% c("CTRL_VEH","CTRL_BAY", "CTRL_BPN", 
                                                                             "FXS_VEH", "FXS_BAY", "FXS_BPN"))


# for (i in 1:length(pathway_overlap_BAY_BPN )) {
#   if ( i %% 10 == 0) {
#     cat("Processing pathway", i, "of", length(pathway_overlap_BAY_BPN ), "\n")
#   }
#   pathway_name <- pathway_overlap_BAY_BPN [i]
# 
#   plot_gsva_boxplot_mutil(gsva_matrix,
#                           condition_list_label = condition_list_label,
#                           pathway_name =  pathway_name,
#                           figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_Overlap_BAY_BPN_all"),
#                           file_name = paste0("GSVA_", pathway_name, "_DRUG"),
#                           fig.height = 6, fig.width = 6)
#   
#     plot_gsva_boxplot_mutil(gsva_matrix,
#                           condition_list_label = condition_list_label_BAY_BPN ,
#                           pathway_name =  pathway_name,
#                           figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_Overlap_BAY_BPN"),
#                           file_name = paste0("GSVA_", pathway_name, "_DRUG"),
#                           fig.height = 6, fig.width = 6)
# 
# 
# }


for (i in 1:2) {
  if ( i %% 10 == 0) {
    cat("Processing pathway", i, "of", length(pathway_overlap_BAY_BPN ), "\n")
  }
  pathway_name <- pathway_overlap_BAY_BPN [i]

  p1<-plot_gsva_boxplot_mutil(gsva_matrix,
                          condition_list_label = condition_list_label,
                          pathway_name =  pathway_name,
                          figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_Overlap_BAY_BPN_all"),
                          file_name = paste0("GSVA_", pathway_name, "_DRUG"),
                          fig.height = 6, fig.width = 6)


  p2<-  plot_gsva_boxplot_mutil(gsva_matrix,
                          condition_list_label = condition_list_label_BAY_BPN ,
                          pathway_name =  pathway_name,
                          figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_Overlap_BAY_BPN"),
                          file_name = paste0("GSVA_", pathway_name, "_DRUG"),
                          fig.height = 6, fig.width = 6)

  
p_comp <- plot_grid(p1, p2, labels = c("A", "B"))
print(p_comp)
}
write.csv(pathway_overlap_BAY_BPN, "./results/03-Mix-effect/pathway_overlap_BAY_BPN.csv", row.names = FALSE)


```
\newpage
**BAY and BP**

```{r BAY_BP, fig.width=12, fig.height=6}

pathway_overlap_BAY_BP <- unique(result_df_BAY$pathway) %>% intersect(unique(result_df_BP$pathway))


condition_list_label_BAY_BP <- condition_list_label %>% filter(group %in% c("CTRL_VEH","CTRL_BAY", "CTRL_BP", 
                                                                             "FXS_VEH", "FXS_BAY", "FXS_BP"))

# 
# for (i in 1:length(pathway_overlap_BAY_BP )) {
#   if ( i %% 10 == 0) {
#     cat("Processing pathway", i, "of", length(pathway_overlap_BAY_BP ), "\n")
#   }
#   pathway_name <- pathway_overlap_BAY_BP [i]
# 
#   plot_gsva_boxplot_mutil(gsva_matrix,
#                           condition_list_label = condition_list_label,
#                           pathway_name =  pathway_name,
#                           figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_Overlap_BAY_BP_all"),
#                           file_name = paste0("GSVA_", pathway_name, "_DRUG"),
#                           fig.height = 6, fig.width = 6)
# 
#     plot_gsva_boxplot_mutil(gsva_matrix,
#                           condition_list_label = condition_list_label_BAY_BP ,
#                           pathway_name =  pathway_name,
#                           figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_Overlap_BAY_BP"),
#                           file_name = paste0("GSVA_", pathway_name, "_DRUG"),
#                           fig.height = 6, fig.width = 6)
# 
# 
# }

for (i in 1:2 ) {
  if ( i %% 10 == 0) {
    cat("Processing pathway", i, "of", length(pathway_overlap_BAY_BP ), "\n")
  }
  pathway_name <- pathway_overlap_BAY_BP[i]
  
  p1<-plot_gsva_boxplot_mutil(gsva_matrix,
                          condition_list_label = condition_list_label,
                          pathway_name =  pathway_name,
                          figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_Overlap_BAY_BP_all"),
                          file_name = paste0("GSVA_", pathway_name, "_DRUG"),
                          fig.height = 6, fig.width = 6)

  p2<-plot_gsva_boxplot_mutil(gsva_matrix,
                          condition_list_label = condition_list_label_BAY_BP ,
                          pathway_name =  pathway_name,
                          figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_Overlap_BAY_BP"),
                          file_name = paste0("GSVA_", pathway_name, "_DRUG"),
                          fig.height = 6, fig.width = 6)
  
  p_comp <- plot_grid(p1, p2, labels = c("A", "B"))
  print(p_comp)
}

write.csv(pathway_overlap_BAY_BP, "./results/03-Mix-effect/pathway_overlap_BAY_BP.csv", row.names = FALSE)



```

\newpage
**BPN and BP**

```{r BPN_BP, fig.width=12, fig.height=6}

pathway_overlap_BPN_BP <- unique(result_df_BPN$pathway) %>% intersect(unique(result_df_BP$pathway))


condition_list_label_BPN_BP <- condition_list_label %>% filter(group %in% c("CTRL_VEH","CTRL_BPN", "CTRL_BP", 
                                                                            "FXS_VEH", "FXS_BPN", "FXS_BP"))

# 
# for (i in 1:length(pathway_overlap_BPN_BP )) {
#   if ( i %% 10 == 0) {
#     cat("Processing pathway", i, "of", length(pathway_overlap_BPN_BP ), "\n")
#   }
#   pathway_name <- pathway_overlap_BPN_BP [i]
# 
#   plot_gsva_boxplot_mutil(gsva_matrix,
#                           condition_list_label = condition_list_label,
#                           pathway_name =  pathway_name,
#                           figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_Overlap_BPN_BP_all"),
#                           file_name = paste0("GSVA_", pathway_name, "_DRUG"),
#                           fig.height = 6, fig.width = 6)
# 
#     plot_gsva_boxplot_mutil(gsva_matrix,
#                           condition_list_label = condition_list_label_BPN_BP ,
#                           pathway_name =  pathway_name,
#                           figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_Overlap_BPN_BP"),
#                           file_name = paste0("GSVA_", pathway_name, "_DRUG"),
#                           fig.height = 6, fig.width = 6)
# 
# 
# }

for (i in 1:3 ) {
  if ( i %% 10 == 0) {
    cat("Processing pathway", i, "of", length(pathway_overlap_BPN_BP ), "\n")
  }
  pathway_name <- pathway_overlap_BPN_BP[i]
  
  p1<-plot_gsva_boxplot_mutil(gsva_matrix,
                          condition_list_label = condition_list_label,
                          pathway_name =  pathway_name,
                          figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_Overlap_BPN_BP_all"),
                          file_name = paste0("GSVA_", pathway_name, "_DRUG"),
                          fig.height = 6, fig.width = 6)

  p2<-plot_gsva_boxplot_mutil(gsva_matrix,
                          condition_list_label = condition_list_label_BPN_BP ,
                          pathway_name =  pathway_name,
                          figure_folder = file.path(result_folder,"03-Mix-effect","Boxplot_Overlap_BPN_BP"),
                          file_name = paste0("GSVA_", pathway_name, "_DRUG"),
                          fig.height = 6, fig.width = 6)
  
p_comp <- plot_grid(p1, p2, labels = c("A", "B"))
print(p_comp)
}

write.csv(pathway_overlap_BPN_BP, "./results/03-Mix-effect/pathway_overlap_BPN_BP.csv", row.names = FALSE)

```



\newpage
# Session information
```{r}
sessionInfo()
```

